# Handwritten Digits Classification Competition

[MNIST](http://yann.lecun.com/exdb/mnist/) is a handwritten digits image data set created by Yann LeCun. Every digit is represented by a 28x28 image. It is the "Hello World!" equivalent in Deep Learning. There's a [long-term hosted competition](https://www.kaggle.com/c/digit-recognizer) on Kaggle using this data set. This example is based on [mxnet](https://github.com/dmlc/mxnet/tree/master/R-package).

### Data Loading

First, let us download the data if it does not already exist. If data is not available at that link, download from [here](https://www.kaggle.com/c/digit-recognizer/data).

```{r, echo=FALSE}
dataDirectory <- "../data"
if (!file.exists(paste(dataDirectory,'/train.csv',sep="")))
{
  link <- 'https://apache-mxnet.s3-accelerate.dualstack.amazonaws.com/R/data/mnist_csv.zip'
  if (!file.exists(paste(dataDirectory,'/mnist_csv.zip',sep="")))
    download.file(link, destfile = paste(dataDirectory,'/mnist_csv.zip',sep=""))
  unzip(paste(dataDirectory,'/mnist_csv.zip',sep=""), exdir = dataDirectory)
  if (file.exists(paste(dataDirectory,'/test.csv',sep="")))
    file.remove(paste(dataDirectory,'/test.csv',sep=""))
}
```

Read data into R and convert to matrices. The data is created in augment.R.

```{r}
require(mxnet)
source("img_ftns.R")
options(scipen=999)


train <- read.csv("../data/train_augment.csv", header=TRUE)
test <- read.csv("../data/test_0.csv", header=TRUE)

train <- train[sample(1:nrow(train)), ] # shuffle

train.y <- train[,1]
test.y <- test[,1]
test <- data.matrix(test)
train <- data.matrix(train)

train.x <- train[,-1]
test <- test[,-1]
```

The train dataset has `r nrow(train)` rows.

Linearly transform it into [0,1] by dividing by 255. We also transpose the input matrix to npixel x nexamples, which is the column major format accepted by mxnet (and the convention of R).

```{r}
train.x <- t(train.x/255)
test <- t(test/255)
```

## LeNet

Create a model based on the LeNet architecture. This is a the same model as in the previous chapter, we have 2 sets of convolutional+pooling layers and then a Flatten layer and finally two Dense layers.

```{r}
mx.set.seed(0)
# input
data <- mx.symbol.Variable('data')
# first conv
conv1 <- mx.symbol.Convolution(data=data, kernel=c(5,5), num_filter=20)
tanh1 <- mx.symbol.Activation(data=conv1, act_type="tanh")
pool1 <- mx.symbol.Pooling(data=tanh1, pool_type="max",
                           kernel=c(2,2), stride=c(2,2))
# second conv
conv2 <- mx.symbol.Convolution(data=pool1, kernel=c(5,5), num_filter=50)
tanh2 <- mx.symbol.Activation(data=conv2, act_type="tanh")
pool2 <- mx.symbol.Pooling(data=tanh2, pool_type="max",
                           kernel=c(2,2), stride=c(2,2))
# first fullc
flatten <- mx.symbol.Flatten(data=pool2)
fc1 <- mx.symbol.FullyConnected(data=flatten, num_hidden=500)
tanh3 <- mx.symbol.Activation(data=fc1, act_type="tanh")
# second fullc
fc2 <- mx.symbol.FullyConnected(data=tanh3, num_hidden=10)
# loss
lenet <- mx.symbol.SoftmaxOutput(data=fc2)

train.array <- train.x
dim(train.array) <- c(28, 28, 1, ncol(train.x))
test.array <- test
dim(test.array) <- c(28, 28, 1, ncol(test))

devices <- mx.gpu()
tic <- proc.time()
model2 <- mx.model.FeedForward.create(lenet, X = train.array, y = train.y,
                                     ctx = devices, num.round = 5,
                                     array.batch.size = 100,
                                     learning.rate = 0.05, momentum = 0.9, wd = 0.00001,
                                     eval.metric = mx.metric.accuracy,
                                     batch.end.callback = mx.callback.log.train.metric(100))
print(proc.time() - tic)
preds2 <- predict(model2, test.array)
pred.label <- max.col(t(preds2)) - 1

res <- data.frame(cbind(test.y,pred.label))
table(res)
accuracy2 <- sum(res$test.y == res$pred.label) / nrow(res)
```

The accuracy of our model with augmented train data is `r accuracy2`.

## Test Time Augmentation (TTA)
```{r}
test_data <- read.csv("../data/test_augment.csv", header=TRUE)

test.y <- test_data[,1]
test <- data.matrix(test_data)
test <- test[,-1]
test <- t(test/255)

test.array <- test
dim(test.array) <- c(28, 28, 1, ncol(test))

preds3 <- predict(model2, test.array)
dfPreds3 <- as.data.frame(t(preds3))
# res is a data frame with our predictions after train data augmentation,
# i.e. 4200 rows
res$pred.label2 <- 0
for (i in 1:nrow(res))
{
  sum_r <- dfPreds3[((i-1)*3)+1,] +
    dfPreds3[((i-1)*3)+2,] + dfPreds3[(i*3),] 
  res[i,"pred.label2"] <- max.col(sum_r)-1
}
table(res[,c("test.y","pred.label2")])
accuracy3 <- sum(res$test.y == res$pred.label2) / nrow(res)
```

The accuracy of our CNN model with augmented train data and Test Time Augmentation (TTA) is `r accuracy3`.

```{r}
tta_incorrect <- nrow(res[res$test.y != res$pred.label2 & res$test.y == res$pred.label,])

tta <- res[res$test.y == res$pred.label2 & res$test.y != res$pred.label,c("pred.label","pred.label2")]
```

Number of rows where Test Time Augmentation (TTA) changed the prediction to the correct value `r nrow(tta)`.

Number of rows where Test Time Augmentation (TTA) changed the prediction to the incorrect value `r tta_incorrect`.

```{r}
tta
table(tta)
```
