
The Fashion MINST dataset is available @ [Kaggle](https://www.kaggle.com/zalando-research/fashionmnist).

Download the data and put it in the data folder. Now we can read data into R and get an idea of how it looks. 

```{r}
train <- read.csv("../data/fashion-mnist_train.csv", header=TRUE, nrows=20)
```

We have `r nrow(train)` rows and `r ncol(train)` columns, we will look at the first 5 columns and the last 5 columns:

```{r}
head(train[,1:5])
head(train[,(ncol(train)-5):ncol(train)])
```

We have 785 columns, the first column is the data label and then we have 784 columns named pixel0,...,pixel783 with the pixel values. Our images are 28x28 = 784, so everything looks ok.

Before we start building models, it is always a good idea to ensure your data is in the correct format and that your features and labels are aligned correctly. Lets plot the first 9 instances with their data labels. To do this, we create a helper function called plotInstance that takes in the pixel values and outputs the image with an optional header.

Each training and test example is assigned to one of the following labels:
0 T-shirt/top
1 Trouser
2 Pullover
3 Dress
4 Coat
5 Sandal
6 Shirt
7 Sneaker
8 Bag
9 Ankle boot 

```{r}
plotInstance <-function (row,title="")
{
  mat <- matrix(row,nrow=28,byrow=TRUE)
  mat <- t(apply(mat, 2, rev))
  image(mat, main = title,axes = FALSE, col = grey(seq(0, 1, length = 256)))
}
labels<-c("T-shirt/top","Trouser","Pullover","Dress","Coat","Sandal","Shirt","Sneaker","Bag","Ankle boot")

par(mfrow = c(4, 4))
par(mar=c(2,2,2,2))
for (i in 1:16)
{
  row <- as.numeric(train[i,2:ncol(train)])
  num_label<-train[i,1]
  plotInstance(row, paste("index:",i,", label =",labels[num_label+1]))
}
```
