# Handwritten Digits Classification Competition

[MNIST](http://yann.lecun.com/exdb/mnist/) is a handwritten digits image data set created by Yann LeCun. Every digit is represented by a 28x28 image. It is the "Hello World!" equivalent in Deep Learning. There's a [long-term hosted competition](https://www.kaggle.com/c/digit-recognizer) on Kaggle using this data set. This example is based on [mxnet](https://github.com/dmlc/mxnet/tree/master/R-package).

### Data Loading

First, let us download the data if it does not already exist. If data is not available at that link, download from [here](https://www.kaggle.com/c/digit-recognizer/data).

```{r, echo=FALSE}
dataDirectory <- "../data"
if (!file.exists(paste(dataDirectory,'/train.csv',sep="")))
{
  link <- 'https://apache-mxnet.s3-accelerate.dualstack.amazonaws.com/R/data/mnist_csv.zip'
  if (!file.exists(paste(dataDirectory,'/mnist_csv.zip',sep="")))
    download.file(link, destfile=paste(dataDirectory,'/mnist_csv.zip',sep=""))
  unzip(paste(dataDirectory,'/mnist_csv.zip',sep=""), exdir=dataDirectory)
  if (file.exists(paste(dataDirectory,'/test.csv',sep="")))
    file.remove(paste(dataDirectory,'/test.csv',sep=""))
}
if (!file.exists(paste(dataDirectory,'/train.csv',sep="")))
{
  library(keras)
  mnist <- dataset_mnist()
  c(c(x_train,y_train),c(x_test,y_test)) %<-% dataset_mnist()
  x_train <- array_reshape(x_train,c(dim(x_train)[1],dim(x_train)[2]*dim(x_train)[3]))
  y_train <- array_reshape(y_train,c(length(y_train),1))
  data_mnist <- as.data.frame(cbind(y_train,x_train))
  dim(data_mnist)
  colnames(data_mnist)[1] <- "label"
  colnames(data_mnist)[2:ncol(data_mnist)] <- paste("pixel",seq(1,784),sep="")
  write.csv(data_mnist,paste(dataDirectory,'/train.csv',sep=""),row.names=FALSE)
}
```

Read data into R and convert to matrices. We want to get an estimate of accuracy on test data, so we split the training data into a train set and test set. Because we have a large number of rows, we can use a split ratio of 90/10.

```{r}
require(mxnet)
options(scipen=999)

dfMnist <- read.csv("../data/train.csv", header=TRUE)
yvars <- dfMnist$label
dfMnist$label <- NULL

set.seed(42)
train <- sample(nrow(dfMnist),0.9*nrow(dfMnist))
test <- setdiff(seq_len(nrow(dfMnist)),train)
train.y <- yvars[train]
test.y <- yvars[test]
train <- data.matrix(dfMnist[train,])
test <- data.matrix(dfMnist[test,])

rm(dfMnist,yvars)
```

Each image is represented as row of 784 (28x28) pixel values. The value of each pixel is in the range 0-255, we linearly transform it into 0-1 by dividing by 255. We also transpose the input matrix to because column major format in order to use it in mxnet.

```{r}
train <- t(train / 255.0)
test <- t(test / 255.0)
```

Before creating a model, we should check that our dataset is balanced, i.e. the number of instances for each digit is reasonably even:

```{r}
table(train.y)
```

### Base model (no convolutional layers)

Now we have the data. The next step is to configure the structure of our first model.

```{r}
data <- mx.symbol.Variable("data")
fullconnect1 <- mx.symbol.FullyConnected(data, name="fullconnect1", num_hidden=256)
activation1  <- mx.symbol.Activation(fullconnect1, name="activation1", act_type="relu")
fullconnect2 <- mx.symbol.FullyConnected(activation1, name="fullconnect2", num_hidden=128)
activation2  <- mx.symbol.Activation(fullconnect2, name="activation2", act_type="relu")
fullconnect3 <- mx.symbol.FullyConnected(activation2, name="fullconnect3", num_hidden=10)
softmax      <- mx.symbol.SoftmaxOutput(fullconnect3, name="softmax")
```

1. In `mxnet`, we use its own data type `symbol` to configure the network. `data <- mx.symbol.Variable("data")` use `data` to represent the input data, i.e. the input layer.
2. Then we set the first hidden layer by `fullconnect1 <- mx.symbol.FullyConnected(data, name="fullconnect1", num_hidden=256)`. This layer has `data` as the input, its name and the number of hidden neurons.
3. The activation is set by `activation1 <- mx.symbol.Activation(fullconnect1, name="relu1", act_type="relu")`. The activation function takes the output from the first hidden layer `fullconnect1`.
4. The second hidden layer takes the result from `activation1` as the input, with its name as "fullconnect2" and the number of hidden neurons as 128.
5. the second activation is almost the same as `activation1`, except we have a different input source and name.
6. This is the output layer. There is 10 categories (digits), so we set the number of neurons to 10.
7. Finally we set the activation to softmax to get a probabilistic prediction.

Now lets train the base model. I have a GPU installed, so I can use that.

```{r}
devices <- mx.gpu()
mx.set.seed(0)
model <- mx.model.FeedForward.create(softmax, X=train, y=train.y,
                                     ctx=devices,array.batch.size=128,
                                     num.round=10,
                                     learning.rate=0.05, momentum=0.9,
                                     eval.metric=mx.metric.accuracy,
                                     epoch.end.callback=mx.callback.log.train.metric(1))
```

### Create predictions for base model

To make prediction, we call the predict function. We can then create a confusion matrix and calculate our accuracy level on test data.

```{r}
preds1 <- predict(model, test)
pred.label1 <- max.col(t(preds1)) - 1
res1 <- data.frame(cbind(test.y,pred.label1))
table(res1)
accuracy1 <- sum(res1$test.y == res1$pred.label1) / nrow(res1)
```

The accuracy of our base model is `r accuracy1`. Not bad, but lets see if we can improve on it.

## LeNet

Now we can create a model based on the LeNet architecture. This is a very simple model, we have 2 sets of convolutional+pooling layers and then a Flatten layer and finally two Dense layers.

```{r}
data <- mx.symbol.Variable('data')
# first convolution layer
convolution1 <- mx.symbol.Convolution(data=data, kernel=c(5,5), num_filter=64)
activation1  <- mx.symbol.Activation(data=convolution1, act_type="tanh")
pool1        <- mx.symbol.Pooling(data=activation1, pool_type="max",
                           kernel=c(2,2), stride=c(2,2))

# second convolution layer
convolution2 <- mx.symbol.Convolution(data=pool1, kernel=c(5,5), num_filter=32)
activation2  <- mx.symbol.Activation(data=convolution2, act_type="relu")
pool2 <- mx.symbol.Pooling(data=activation2, pool_type="max",
                           kernel=c(2,2), stride=c(2,2))

# flatten layer and then fully connected layers
flatten      <- mx.symbol.Flatten(data=pool2)
fullconnect1 <- mx.symbol.FullyConnected(data=flatten, num_hidden=512)
activation3  <- mx.symbol.Activation(data=fullconnect1, act_type="relu")
fullconnect2 <- mx.symbol.FullyConnected(data=activation3, num_hidden=10)
# final softmax layer
softmax <- mx.symbol.SoftmaxOutput(data=fullconnect2)

train.array <- train
dim(train.array) <- c(28,28,1,ncol(train))
test.array <- test
dim(test.array) <- c(28,28,1,ncol(test))

devices <- mx.gpu()
mx.set.seed(0)
model2 <- mx.model.FeedForward.create(softmax, X=train.array, y=train.y,
                                     ctx=devices,array.batch.size=128,
                                     num.round=10,
                                     learning.rate=0.05, momentum=0.9, wd=0.00001,
                                     eval.metric=mx.metric.accuracy,
                                     epoch.end.callback=mx.callback.log.train.metric(1))

# evaluate model
preds2 <- predict(model2, test.array)
pred.label2 <- max.col(t(preds2)) - 1
res2 <- data.frame(cbind(test.y,pred.label2))
table(res2)
accuracy2 <- sum(res2$test.y == res2$pred.label2) / nrow(res2)
```

The accuracy of our CNN model is `r accuracy2`, which is quite an improvement over the accuracy of our base model which was `r accuracy1`.

Finally, we can visualise our model in R.

```{r fig.width=12}
graph.viz(model2$symbol,type="vis")
```

